On The Impact of Atoms of Confusion in JavaScript Code

### 2021-04-29 1_Introduction.tex
minor changes (introduction)

Source code misunderstandings can occur due to many reasons. For instance, a poorly understood requirement might lead to poorly implemented code.  Besides that, programming languages can also have specific constructs that may lead to misunderstandings. In this context, the same applies to \emph{atoms of confusion}---that is, small pieces of code that hinder developers' understanding of the source code of a program for which there are functionally equivalent alternatives that are easier to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17, DBLP:conf/msr/GopsteinZFC18}.

Recent work~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/GopsteinIYDZYC17,Langhout:2021:ACJ} has explored the impact of atoms of confusion on program understanding, and confirmed they negatively impact source code readability. For instance, consider the pair of code snippets in Listing~\ref{fig:lst01} (adapted from \textsc{NervJS/taro}). The code at the left-hand side contains two atoms of confusion discussed in previous work focusing on the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic as Control Flow} and \emph{Post-increment}. At the right-hand side, we present a clean version of the code, that is, without atoms of confusion. According to the experiments of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, removing the atoms \emph{Logic as Control Flow} and \emph{Post-increment} improved the accuracy of program understanding tasks in 41\% and 34\%, respectively.

@@

However, these previous studies focus exclusively on statically typed languages (e.g., \clang, \cpplang, and Java). This way, there is a knowledge gap with respect to the impact of such atoms in dynamic languages such as JavaScript. In this sense, because JavaScript's programming culture differs from that of languages such as \clang and \cpplang, particularly due to the JavaScript's dynamic capabilities and weaker type system, we may find that idioms and constructs that are atoms of confusion in those languages are not particularly difficult to understand for JavaScript programmers and vice-versa. Also, besides being the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}, JavaScript contains features that combined make the source code fertile ground for errors~\cite{BugsJS:ICST,ClientSideBugsJS:TSE} and confusion~\cite{JSGoodParts:book}, such as weakly typed language~\cite{DynamicBehaviorJS:PLDI}, object-oriented, functional, and imperative constructs~\cite{JS-CodeSmells:SCAM}, and features such as run-time evaluation~\cite{BugsJS:ICST,ClientSideBugsJS:TSE}, and asynchronous operations.

In this way, to better understand the impact of atoms of confusion in JavaScript, in this paper we present a mixed-method research effort based on a survey, on interviews, and on a mining software repositories study. First, we surveyed 140 JavaScript developers. In the survey, their task was to predict the output of short code snippets, with half of them containing atoms of confusion. The number of correct answers for the \emph{clean version} of the code snippets for 7 out of \na atom candidates is at least 15\% higher than the corresponding \emph{confusing version} of the code snippet (i.e., with an atom candidate). The presence of an atom candidate might also affect the time necessary for submitting a correct answer. For instance, the presence of the Comma Operator atom increases the average time to predict the output by 76.23\%, though the outputs of snippets containing the Pre-Increment atom were correctly predicted 38.19\% faster.

To learn more about the misunderstandings in the snippets used in the survey, we also interviewed 15 experienced professional developers. Participants considered the code snippets without the atom candidates easier to understand in 70\% of the cases, therefore corroborating the survey findings. The participants also discussed other idioms that might lead developers to misunderstand JavaScript code. The occurrence of atoms in open-source repositories might reveal cases where developers should be more or less careful when writing code---and open the possibility to implement new program transformations for cleaning up confusing code. To identify such cases, we mined popular JavaScript projects from GitHub and found four atom candidates that are highly used in JavaScript code. The Implicit Predicate atom, for instance, occurs 19.89 times per 1,000 lines of code.

@@

    \item A list of additional atoms of confusion addressing particular 
    JavaScript constructs and idioms;

>>>>

Source code misunderstandings can occur due to many reasons; for instance, the wrong choice of names for arguments, variables, and methods might hinder developers from understanding a piece of code~\cite{avidan:icpc2017}. Besides that, programming languages can also have specific idioms and constructs that may lead to misunderstandings. That is the case of \emph{atoms of confusion}---small pieces of code that hinder developers' understanding of the source code of a program for which there are functionally equivalent alternatives that are easier to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17, DBLP:conf/msr/GopsteinZFC18}.

Recent work~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/GopsteinIYDZYC17,Langhout:2021:ACJ} has explored the impact of atoms of confusion on program understanding, and confirmed they negatively impact source code readability. For instance, consider the pair of JavaScript code snippets in Listing~\ref{fig:lst01} (adapted from \textsc{NervJS/taro}). The code at the left-hand side contains two atoms of confusion discussed in previous work focusing on the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: \emph{Logic as Control Flow} and \emph{Post-increment}. At the right-hand side, we present a clean version of the code, that is, without atoms of confusion. According to the experiments of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, removing the atoms \emph{Logic as Control Flow} and \emph{Post-increment} improved the accuracy of program understanding tasks in 41\% and 34\%, respectively.

@@

However, these previous studies focus exclusively on statically typed languages (e.g., \clang, \cpplang, and Java). This way, there is a knowledge gap with respect to the impact of such atoms in dynamic languages such as JavaScript. In this sense, because JavaScript's programming culture differs from that of languages such as \clang and \cpplang, particularly due to the JavaScript's dynamic capabilities and weaker type system, we may find that idioms and constructs that are atoms of confusion in those languages are not particularly difficult to understand for JavaScript programmers and vice-versa. Also, besides being the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}, JavaScript contains features that combined make the source code fertile ground for errors~\cite{BugsJS:ICST,ClientSideBugsJS:TSE} and confusion~\cite{JSGoodParts:book}.
%, such as weakly typed language~\cite{DynamicBehaviorJS:PLDI}, object-oriented, functional,
% and imperative constructs~\cite{JS-CodeSmells:SCAM}, and features such as run-time
% evaluation~\cite{BugsJS:ICST,ClientSideBugsJS:TSE}, and asynchronous operations.

In this way, to better understand the impact of atoms of confusion in JavaScript, in this paper we present a mixed-method research effort based on a survey, on interviews, and on a mining software repositories study. First, we surveyed 140 JavaScript developers. In the survey, their task was to predict the output of small code snippets, with half of them containing atoms of confusion. The number of correct answers for the \emph{clean version} of the code snippets for 7 out of \na atom candidates is at least 15\% higher than the corresponding \emph{confusing version} of the code snippet (i.e., with an atom candidate). The presence of an atom candidate might also affect the time necessary for submitting a correct answer. For instance, the presence of the Comma Operator atom increases the average time to predict the output by 76.23\%, though the outputs of snippets containing the Pre-Increment atom were correctly predicted 38.19\% faster.
To learn more about the misunderstandings in the snippets used in the survey, we also interviewed 15 experienced professional developers. Participants prefered the code snippets without the atom candidates in 70\% of the cases, therefore corroborating the survey findings. The participants also discussed other idioms and constructs that might lead developers to misunderstand JavaScript code.

The frequency in which the atoms appear \emph{in the wild} might reveal cases where developers should be more or less careful when writing JavaScript code. To understand the incidence of atoms, we mined popular JavaScript projects from GitHub and found that atoms of confusion frequently appear in practice. Indeed, four atom candidates are highly used in JavaScript code---the Implicit Predicate atom, for instance, occurs 19.89 times per 1,000 lines of code.

@@

  \item A list of additional, JavaScript-specific
    constructs and idioms that might introduce code
    misunderstanding;
